---
title: "651 class"
author: "Siyu Zou"
date: "2023-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
tq<-qexp(seq(0.01,0.99,length=100), 1/5)
plot(tq)

union(x,y) #求并集

intersect(x,y) #求交集

setdiff(x,y) #求属于x而不属于y的所有元素

setequal(x,y) #判断x与y是否相等

a %in% y #判断a是否为y中的元素

choose(n, k) #n个里面取k个的组合数

combn(x,n) #x中的元素每次取n个的所有组合

combn(x,n,f) #将这些组合用于指定函数f
```

# Lecture 1 &2
## Probability calculus
```{r birthday problem}
n=1:364
pn=n
for (i in 1:364) {
  pn[i]<- 1-prod(365:(365-i+1))/365^i
  }

pn[23]  # 50%

pn[57]  # 99%
```

```{r ploting birthday problem}
plot(
  n[1:100],
  pn[1:100],
  type = "p",
  pch = 19, 
  col = "red",
  lwd = 3,
  xlab = "Number of people in the room",
  ylab = "Probability of at least two same-day birthdays",
  cex.lab = 1.5, cex.axis = 1.5, col.axis = "black"
)


```


### sample
```{r Monte Holl problem}

# help on sample
?sample
# a random permutation
x <- sample (1:6)  #  the default for size == the first argument, so that sample(x) generates a random permutation of the elements of x (or 1:x).

# sampling with replacement
x <- sample ( 1:6, 10, replace=T )  # size = 10

# how many are equal to 3?
sum ( x == 3 )

# nonparametric bootstrap
x <- sample ( 1:10, 10, replace=T )

# 100 Bernoulli trials
sample(c(0,1), 100, replace = TRUE)

```

### Can you write an R code to simulate the Monte Hall problem?
the answer is in Lab 2 notes


## Probability mass function (PMF) for discrete variable

### Bernoulli
```{r Bernoulli example}
?rbinom
# Let X be the result of a single coin flip where X = 0 represents tails and X = 1 represents heads

# for x = 1, 1个硬币(size), 抛20次(n)，每次概率0.5 (p)
rbinom(20,1,0.5)

rbinom(20, 1, 0.3)

rbinom(20,1,seq(0,1,length=20))
seq(0,1,length=20)

```

### Poisson
```{r Poisson example}
x <- 1:100
lambda = 20
plot(x, dpois(x, lambda),type="p",pch=19,
col="blue",lwd=3, xlab="Number of patients",
ylab="Probability",cex.lab=1.5,cex.axis=1.5,
col.axis="blue")


?dpois

# sampling the Poisson distribution 
# A sample of 15 independent days with average number of patients = 20
rpois(15, lambda=20)

# Reconstructing PMF from samples
y <- rpois(10000, lambda=20)
py=rep(0,100)
for (i in 1:100) {
  py[i] <- sum(y==i)/10000 
}

# Add lines to the PMF plot of the Poisson distribution
plot(x, dpois(x, lambda),type="p",pch=19,
col="blue",lwd=3, xlab="Number of patients",
ylab="Probability",cex.lab=1.5,cex.axis=1.5,
col.axis="blue") + 
lines(1:100,py,col="red",lwd=3)

```


## Probability density function for continuous variable
What’s the probability that a randomly selected person from this distribution survives more than 6 years?
f(x) = (e^-x/5)/5
P(X>=6) = $\int_{0}^{Inf}e^-t/5 /5 dt$

```{r exponential}
pexp(6, 1/5, lower.tail = FALSE)
x <- 1:100
plot (dexp(x,1/5 ))
pexp(6, 1/5) - pexp(5, 1/5)
```

Double exponential
```{r double exponential}
x <- seq(-5,5,length(101))
fx <- exp(-abs(x))/2
plot(x, fx, type = "l", col = "lightblue", lwd = 3,
     xlab = "error", ylab = "PDF", cex.lab = 1.5,
     cex.axis = 1.5, col.axis = "blue"
     )
```
consider a more general version of f(.)
f(x) = $1/2σ e^(-|x-μ|/σ)$
is this a valid density?
Plotting double exponentials with mean μ = 2 and σ = 1, 2, 4
```{r double exponentials μ = 2}
# construct the empirical pdf from 100 simulated observation
mu <- 2
sigma = 2
# install.packages("vgam")
# library("vgam")
y = replace(100, mu, sigma)
hist(y, breaks = 20, probability = TRUE, xlab = "error")

# Super impose the theoretical pdf
x=seq(mu-7,mu+7,length=101)
fx=exp(-abs(x-mu)/sigma)/(2*sigma)
plot(x,fx,col="orange",lwd=3)

```

# CDF Cumulative distribution function

# Quantile
What is the 25th percentile of the exponential survival distribution considered before??
• We want to solve (for x)
.25 = F(x)
= 1 − e−x/5
resulting in the solution x = −log(.75) × 5 ≈ 1.44
• Therefore, 25% of the subjects from this population live less than 1.44 years
```{r quantile}
# R can approximate exponential quantiles for you
qexp( p = 0.25, 1/5)  # quartile
qexp( p = 1/3, 1/5)  # tertile
```


# QQ plot
Calculate the quantiles of an exp(5) distribution
```{r qq}
tq<-qexp(seq(0.01,0.99,length=100), 1/5)
plot(tq)
```

How do they compare to the empirical quantiles?
```{r }
x1<-rexp(30, 1/5)
x2<-rexp(30, 1/5)
x3<-rexp(30, 1/5)
eq1<-quantile(x1,seq(0.01,0.99,length=100))
eq2<-quantile(x2,seq(0.01,0.99,length=100))
eq3<-quantile(x3,seq(0.01,0.99,length=100))
```

Plot theoretical versus empirical quantiles (QQ plots)
```{r qq plot}
library(ggplot2)
data <- data.frame( eq1, eq2, eq3, tq )

library(tidyverse)
data %>%
  ggplot() +
  geom_point(aes(tq, eq1 )) +
  geom_point(aes(tq,eq2),method = "lm" ) +
  geom_point(aes(tq, eq3), method = "lm") +
  labs(x = "theoretical Quantiles", y = "Sample Quantiles")

y <- rt(200, df = 5)
qqnorm(y, main = "Normal Q-Q Plot",
       xlab = "Theoretical Quantiles", ylab = "Sample Quantiles",
       plot.it = TRUE, datax = FALSE, )
```


# Lecture 3

## Expected values
E[X] =sum xp(x)
E[g(X)] =Sum g(x)p(x)
Suppose that a die is tossed and X is the number face up
• What is the expected value of X?
```{r expected value}
# Simulate the mean of n die rolls
mx5 <- rep(0, 1000)
mx10 <- rep(0, 1000) 
mx20 <- rep(0, 1000)
mx100 <- rep(0,1000)

for (i in 1:1000) {
  mx5[i] <- mean(sample(1:6, 5, replace = TRUE)) # 5 die rolls
  mx10[i] <- mean(sample(1:6, 10, replace = T))  # 10 die rolls
  mx20[i] <- mean(sample(1:6, 20, replace = T))  # 20 die rolls
  mx100[i] <- mean(sample(1:6, 100, replace = T)) # 100 die rolls
}


die_roll <- data.frame(mx5, mx10, mx20, mx100)
ggplot(die_roll) +
  geom_histogram(aes(mx5))

ggplot(die_roll) +
  geom_histogram(aes(mx100))
```

### continuous rndom variable with pdf f(x),
Expected value = $ \int_x xf(x) dx $

```{r uniform distribution}
y1<-mean(runif(10000))
y2<-mean(runif(10000,0,10))
y3<-mean(runif(10000,-2,15))
```

```{r Gamma distribution}


```

# Lecture 6
Normal distribution
```{r normal}
bx=c(5,2,3) # this is the data
mu=seq(0,6,length=201) # parameter values
likep=rep(0,201)
for (i in 1:201)
febx2=-sum((bx-mu[i])^2)/2
likep[i]=exp(ebx2)/((2*pi)^(length(bx)/2))g
plot(mu,likep,type="l",col="blue",lwd=3)
mle<-mu[which.max(likep)]

```


```{r lecture 6}
theta=seq(1,10,by=0.1)
like=1/theta^3*(theta>=5)
plot(theta,like,type="l",col="blue",lwd=3)


like[theta==6]/like[theta==5]
like[theta==6]/like[theta==4]
theta[which.max(like)] # maximum likelihood
liken=like/max(like)
plot(theta,liken,type="l",col="blue",lwd=3)



```



