---
title: "homework"
author: "Siyu Zou"
date: "2023-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r homework 1 Q6}
set.seed(19990318)
U <- runif(1000, 0, 1)
gamma = 2
beta = 5 
# inverse CDF F^{-1}
x = (-beta*log(1-U))^(1/gamma)
hist(U,main = "Uniform")
hist (x, main =" Simulated Weibull ")

plot(dweibull(1000, 2, 5 ))
# check if X follows a Weibull distribution using a QQ plot
qqplot(qweibull(x, gamma, beta), x, ylab = "simulated", xlab = "theoretical", main = "Weibull QQ plot")
qqline (x, distribution = function (p) qweibull (p,gamma , beta ),lty = 2, col = 2, lwd = 3)

```

## Q4 sensitivity 
Create a plot of the sensitivity by (1 - specificity), which is the true positive rate versus the false positive rate for all of the cut-offs between 0 and 30. This is called an ROC curve.
```{r homework 2 Q4}
library(ggplot2)
library(tidyverse)
D <- 16
ND <- 46
class(D)
T <- c(2, 3, 10, 24, 43, 62 )
NT <- c(60,59, 52, 38,  19 , 0)

TP <- c(2, 3, 7, 12, 15, 16)
TN <- c(46, 46, 43, 34, 18, 0)
sensitivity <- TP/D
specificity <- TN/ND

plot( 1-specificity, sensitivity, main = "ROC curve", type = "b") 

 1-specificity
# data <- data.frame(list(TP, TN,sensitivity,specificity  ))
# data %>%
#   ggplot(aes(1-specificity, sensitivity)) +
#   geom_point() +
#   geom_smooth( method = "gam", se = FALSE ) +
#   coord_cartesian(xlim = c(0, 1), ylim = c(0,1))


prevalence <- seq(0, 1, by = 0.001)
sensitivity <- 0.75
specificity <- 34/(12+34)
PPV <- (sensitivity*prevalence)/( sensitivity*prevalence + (1-specificity)*(1-prevalence) )

plot(prevalence, PPV)

library(riskyr)
plot_curve (
  sens = 0.7500,
  spec =1-0.26086957,
  prev = NA
)


```


## Q5

```{r Q5}
# Simulate the mean of different n keys 
key_no_replace <- rep(0, 1000) # 1000 keys
key_replace <- rep(0, 1000)

Key_rep <- 10000 # how many repeats you want to do

# First, sampling without replacement
for (i in 1:Key_rep){
  # simulate true key number
  key_no_replace[i] <- sample(x = key, size = 1, replace = FALSE)
}

mean( key_no_replace )

# Second, sampling with replacement
for (i in 1:Key_rep){
  # simulate true key number
  key_replace[i] <- sample(x = key, size = 1, replace = T)
}
summary(key_replace)

```
### Without replacement
```{r Q5 simulating}
# without replacement
ntrail = 1000
results = vector()

for (i in 1:ntrail) {
  select_key = sample(1:1000, 1, replace = FALSE)
  right_key = sample(1:1000, 1, replace = FALSE)
  select_list = c()
  count = 1
  all_keys = 1:1000
  while (select_key != right_key) {
    count = count  + 1
    select_list = append(select_list, select_key)
    remaining_keys = setdiff(all_keys, c(select_list, select_key))
    if (length(remaining_keys) == 0) {
      break
    }
    select_key = sample(remaining_keys, 1, replace = FALSE)
  }
  results[i] = count
}

mean(results)
results
```

### With replacement
```{r with replacement}
ntrail = 1000
results = vector()
  
for (i in 1:ntrail) {
  select_key = sample(1:1000, 1, replace = TRUE)
  right_key = sample(1:1000, 1, replace = TRUE)
  count = 1
  all_keys = 1000
  while (select_key != right_key) {
    count = count  + 1
    select_key = sample(1:1000, 1, replace = TRUE)
    right_key = sample(1:1000, 1, replace = TRUE)
  }
  results[i] = count
}

mean(results)
```

## Q10
In R produces 1,000 averages of 10 die rolls. That is, it’s like taking ten dice, rolling them, averaging the results and repeating this 1,000 times
```{r Q10}
set.seed(19980415)
temp <- matrix(sample(1 : 6, 1000 * 10, replace = TRUE), 1000)
xBar <- apply(temp, 1, mean)
hist(xBar, xlab="Sample Average", main="Average of 10 die rolls")

```

b. The expected value of sample averages of iid random variables is approximately equal to the expectation for a single die roll. Thus, we expect the mean to be close to 3.5 as this is the expectation for a single die roll.

```{r Q10 b}
mean(xBar)

```
c. Since the variance of a single die roll is 2.92, we expect the variance of the sample average of 10 dice rolls to be 2.92/100 , thus the standard deviation of these sample means should be approximately sqrt(2.92/10) = 0.54

sd(xBar)



## Q11
Problem 11. (Practice)
You are playing a game with a friend where you flip a coin and if it comes up heads you give her a dollar and if it comes up tails she gives you a dollar. You play the game ten times.
a. What is the expected total earnings for you? (Show your work; state your assumptions.)
b. What is the variance of your total earnings? (Show your work; state your assumptions.)
c. Suppose that the coin is biased and you have a .4 chance of winning for each flip. Repeat the calculations in parts a and b
d. Suppose each of you has $10 when you start the game. The game is played indefinitely until one person loses all of their money. Write R code to simulate this game. Calculate the expected duration of the game, i.e. on average how long (how many flips) will the game last?

```{r Q11}
results_list = list() ##list initialization
for (p in seq(.1,.9,.1)){
  ntry = 1000
  results = vector()
  for (i in 1:ntry){
    a = 10
    b = 10
    count = 0
    while (a>0 & b>0){
      flip = rbinom(1, 1, p)
      if (flip == 1){a=a+1;b=b-1}
      if (flip == 0){a=a-1;b=b+1}
      count = count + 1
      }
    results[i] = count
  }
results_list[[toString(p)]] = results
}

library(reshape2)
results_df = do.call(cbind.data.frame,results_list)
results_df %>% 
  melt(id.vars=NULL) %>% 
  ggplot(aes(x=value)) +
  geom_histogram(bins = 15) +
  facet_wrap(vars(variable),nrow = 3,scales = "free_x") +
  theme_bw(base_size = 12)+
  labs(x='Expected duration of the game')

apply(results_df,2,summary)
```


## Q12
Note that the code xBar <- apply(matrix(runif (1000 ∗ 10), 1000), 1, mean) produces 1,000 averages of 10 uniforms.
```{r Q12}
xBar <- apply(matrix(runif (1000 * 10), 1000), 1, mean)
xBar
runif (1000 * 10)
matrix(runif (1000 * 10), 1000)   # 1000 row * 10 col, nrow = 1000
hist(xBar, xlab="Sample Average", main="Average of 10 Uniform RV")
```





