---
title: "651 homework"
author: "Siyu Zou"
date: "2023-09-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 3
## Q1 plot the likelihood associated with the experiment, renormalize the likelihood so that its maximum is one.
```{r}
p <- seq(0,1,by=0.01)
likelihood <- choose(10,7)*p^7*(1-p)^3
likelihood.norm <- likelihood/max(likelihood) 
plot( p, likelihood.norm, type = "b"  )

choose(10,7)*0.5^10 + choose(10,8)*0.5^10 + choose(10,9)*0.5^10 + choose(10,10)*0.5^10

p <- seq(0,1,by=0.01)
likelihood_2 <- choose(9,2)*(1-p)^3*p^7
likelihood.norm2 <- likelihood_2/max(likelihood_2) 
plot( p, likelihood.norm2, type = "b"  )

1- ( 0.5^3 + choose(3,2)*0.5^4 + choose(4,2)*0.5^5 + choose(5,2)*0.5^6 + choose(6,2)*0.5^7 + choose(7,2)*0.5^8 + choose(8,2)*0.5^9 )

```

```{r Q2}
#a. 5/6
pnorm(5/6)

#b. What DBP represents the 90th, 95th and 97.5th percentiles of normal distribution
qnorm(0.90)
12*qnorm(0.90) + 80
qnorm(0.95)
12*qnorm(0.95) + 80
qnorm(0.975)
12*qnorm(0.975) + 80

# c. 
pnorm(1, lower.tail = FALSE)
pnorm(2, lower.tail = FALSE)
pnorm(3, lower.tail = FALSE)

# d Suppose that 10 people are sampled from this population. What’s the probability that 50% (5) of them have a SBP larger than 140?
pnorm(20/11)
pnorm(20/11, lower.tail = FALSE)

choose(10,5)*(pnorm(20/11))^5 * (1-pnorm(20/11))^5
choose(10,5)*0.9655^5 * 0.0345^5
252 * 0.8389989 * 0.0345^5

choose(1000,500)*0.9655^500 * 0.0345^500


# f 
0.0345* (1-0.7976)

# g
(81.3-80)/12*sqrt(200)
pnorm(1.532065)

(81.3-120)/11*sqrt(200)
pnorm((81.3-120)/11*sqrt(200))

```


```{r Q3}
rexp(1000, 1)
mean(rexp(1000, 1))
var(rexp(1000, 1))

# x <- seq(0,5,length = 1000)
# y <- mean(rexp(1000,1))
# n <- 1000
```

# Retain the same 1,000 observations from part a. Plot the sequential sample means by observation number. Explain the resulting plot.
```{r}
# example
x1=rbinom(100,1,0.5)
x2=rbinom(100,1,0.5)
x3=rbinom(100,1,0.5)
xbar1=cumsum(x1)/(1:100)
xbar2=cumsum(x2)/(1:100)
xbar3=cumsum(x3)/(1:100)
plot(xbar1)

set.seed(19990318)
sample_size <- 1000
sample_data <- rexp(sample_size, 1)
sequence_mean <- cumsum(sample_data)/(1:1000)
plot(1:sample_size, sequence_mean, type = "l", col = "blue",
     xlab = "Observation Number", ylab = "Sequential Sample Mean",
     main = "Sequential Sample Means")


## Same like above
sample_size <- 1000
sample_data <- rexp(sample_size, 1)

# Initialize a vector to store sample means
sample_means <- sequence( 1000 ,0, 0)
sequential_means <- numeric(sample_size)
total <- 0
# Calculate the sequential sample means
for (i in 1:1000) {
  total <- total + sample_data[i]
  sample_means[i] <- total / i
  sequential_means[i] <- sample_means[i]
}
# Create a plot of sequential sample means by observation number
plot(1:sample_size, sequential_means, type = "l", col = "blue",
     xlab = "Observation Number", ylab = "Sequential Sample Mean",
     main = "Sequential Sample Means")

# Add a horizontal line at the population mean for reference
abline(h = 1, col = "red", lty = 2)

# Add a legend
legend("topright", legend = c("Sample Means", "Population Mean"),
       col = c("blue", "red"), lty = c(1, 2))

```


Plot a histogram of the 1, 000 numbers. Does it look like a exponential density?
```{r c}
hist(sequence_mean, prob = T, breaks = 100)
```

d. Now sample 1,000 sample means from this distribution, each comprised of 100 observations. What numbers should the average and variance of these 1,000 numbers be equal to and why?

```{r Q3 d e}
sample_data <- numeric(1000)

for (i in 1:1000) {
  sample_data[i] = mean(rexp(100,1))
}

mean <- mean(sample_data)
variance <- var(sample_data)

# plot(1:1000, mean)
hist( sample_data, prob = T, breaks = 35 )
```

Now sample 1,000 sample variances from this distribution, each comprised of 100
observations. Take the average of these 1, 000 variances. What property does
this illustrate and how?
```{r Q3 f}
sample_variance <- numeric(1000)

for (i in 1:1000) {
  sample_variance[i] = var(rexp(100,1))
}

mean <- mean(sample_variance)
variance <- var(sample_variance)

# plot(1:1000, mean)
hist( sample_variance, prob = T, breaks = 35 )
```

```{r Q5 d}
library(ggplot2)
beta <- seq(0, 5, by = 0.1)
data <- c(2.0, 1.5, 1.0, 1.2, 1.3, 4.4, 1.0, 4.9, 1.5, 1.2) # observed values
likelihood <- beta^10 / ( cumprod(data)[10]^(beta+1)) 

likelihood.norm <- likelihood/max(likelihood) # easier to interpret

ggplot(data.frame(beta, likelihood.norm), aes(x = beta, y = likelihood.norm)) +
  geom_line() + 
  labs(title = "Likelihood of Pareto Distribution", x = "Beta", y = "Normalized Likelihood") + 
  theme_bw()

# Identify reference lines
# r8 <- beta[range(which(likelihood.norm > 1/8))] 
# r16 <- beta[range(which(likelihood.norm > 1/16))]

#   geom_vline(xintercept = r8, color = 'red', linetype = 'dashed')+
#   geom_vline(xintercept = r16, color = 'blue', linetype = 'dashed') +


```




